Analysis of transit delays in the Toronto Transit Commission: 2024

### 2024-09-26 paper/paper.qmd
Update paper.qmd

# Introduction

Having operated for over 100 years, the Toronto Transit Commission (TTC) has become one of the most visible and representative public transportation organizations in North America. It is well-known as the primary mode of public transportation in Toronto's daily life. Within the TTC, the bus network typically experiences higher daily ridership compared to streetcars, while subways accommodate the largest volume of passengers due to their central role in covering longer routes. Frequent riders often notice delays that arise on a daily basis. As delays become worse and services becoming increasingly inconsistent, it is important to examine the potential causes, allowing the TTC to identify and implement effective solutions.

@@

# Data Analysis
The dataset provides information that offers insights of public transit delays across different types of transportation.

@@

# Data Visualization
```{r}

Figure 1: Distribution of Bus and Streetcar Delay Reasons

@@

```{r}

@@

```{r}

Figure 3:The trend of delays based on the day of the week

>>>>

# Introduction
Having operated for over 100 years, the Toronto Transit Commission (TTC) has become one of the most visible and representative public transportation organizations in North America. It is well-known as the primary mode of public transportation in Toronto's daily life. Within the TTC, the bus network typically experiences higher daily ridership compared to streetcars, while subways accommodate the largest volume of passengers due to their central role in covering longer routes. Frequent riders often notice delays that arise on a daily basis. As delays become worse and services becoming increasingly inconsistent, it is important to examine the potential causes, allowing the TTC to identify and implement effective solutions.

@@

# Data Analysis
The dataset provides information that offers insights of public transit delays across different types of transportation. The y-axis represents  The x-axis corresponds ,representing the total income for each year.

@@

# Data Visualization
```{r}
# Load the required libraries
library(ggplot2)
library(dplyr)

# Read the data from both CSV files
bus_data <- read.csv("ttc_bus_delay_data_2024.csv")
streetcar_data <- read.csv("ttc_streetcar_delay_data_2024.csv")

# Combine the two data frames into one
combined_data <- bind_rows(bus_data, streetcar_data)

# Summarize the data by incident type (combining bus and streetcar delays)
incident_summary <- combined_data %>%
  group_by(Incident) %>%
  summarise(Total_Incidents = n()) %>%
  arrange(desc(Total_Incidents))

# Plot the total incidents for each incident type using ggplot2
ggplot(incident_summary, aes(x = reorder(Incident, -Total_Incidents), y = Total_Incidents)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Reasons of Bus and Streetcar Delay", 
       x = "Incident Type", 
       y = "Total Incidents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for better readability

```
Figure 1: Distribution of Bus and Streetcar Delay Reasons

@@

```{r}
library(ggplot2)
library(dplyr)

subway_data <- read.csv("Subway_count_2024.csv")

subway_data_clean <- subway_data %>%
  filter(!is.na(Delay.Group) & !is.na(Count) & Count > 0)

# Summarize the data by Delay.Group
delay_group_summary <- subway_data_clean %>%
  group_by(Delay.Group) %>%
  summarise(Total_Count = sum(Count)) %>%  # Sum up the Count column for each Delay.Group
  arrange(desc(Total_Count))  # Arrange by total delays in descending order

# Plot the total incidents for each delay group using ggplot2
ggplot(delay_group_summary, aes(x = reorder(Delay.Group, -Total_Count), y = Total_Count)) +
  geom_bar(stat = "identity", fill = "tomato") +
  labs(title = "Total Delays by Delay Group (TTC Subway)", 
       x = "Delay Group", 
       y = "Total Delays") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for better readability

@@

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(lubridate)

# Define file paths (replace with actual paths)
file1 <- ("ttc_subway_delay_data_2024.csv")
file2 <- ("ttc_bus_delay_data_2024.csv")
file3 <- ("ttc_streetcar_delay_data_2024.csv")

data1 <- read.csv(file1)
data2 <- read.csv(file2)
data3 <- read.csv(file3)

# Inspect column names and structure
print(names(data1))
print(names(data2))
print(names(data3))

# Assuming the date column might be named 'Date'
data1$date <- as.Date(data1$Date, format = "%Y-%m-%d")
data2$date <- as.Date(data2$Date, format = "%Y-%m-%d")
data3$date <- as.Date(data3$Date, format = "%Y-%m-%d")

# Extract the day of the week for each file and count incidents
day_counts_file1 <- data1 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 1")

day_counts_file2 <- data2 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 2")

day_counts_file3 <- data3 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 3")

# Combine the results into a single dataframe
combined_day_counts <- bind_rows(day_counts_file1, day_counts_file2, day_counts_file3)

# Plot the variation of delay incidents by day of the week for each file
ggplot(combined_day_counts, aes(x = day_of_week, y = total_count, color = file, group = file)) +
  geom_line() +
  geom_point() +
  labs(title = "Variation of Delay Incidents by Day of the Week",
       x = "Day of the Week",
       y = "Total Count",
       color = "File") +
  theme_minimal()
```
Figure 3:The trend of delays based on the day of the week
